---
title: "Hexoskin"
author: "Jonathan Slaney and Daniel Fuller"
date: "03/10/2019"
output: html_document
---

## Necessary packages

```{r}
library(knitr)
library(tidyverse)
library(tidymodels)
library(ggmap)
library(lubridate)
library(readr)
library(psych)
library(ggpubr)
library(ggpmisc)
library(effsize)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Looping over data and creating city and participant columns

```{r}
hexodata_temp <- NULL

hexodata_temp <- dir(path = "raw/", full.names = TRUE, pattern = "*.csv", recursive = TRUE) %>%
   purrr::map(function(i){
     dfhx <- read.csv(i, header = TRUE)
     dfhx$city <- stringr::str_sub(i, 6, 10)
     dfhx$participants <- stringr::str_sub(i, 16, 17)
     hexodata_temp <- bind_rows(hexodata_temp, dfhx)
   })

hexo_data_temp <- bind_rows(hexodata_temp)

glimpse(hexo_data_temp)

table(hexo_data_temp$participants)
table(hexo_data_temp$city)
```

### Change column names, create hexo_data and write to csv

```{r}
hexo_data <- hexo_data_temp %>%
  dplyr::rename(
     time = time..s.256., 
     respiratory_rate = breathing_rate..rpm...api.datatype.33.., 
     heart_rate = heart_rate..bpm...api.datatype.19.., 
     minute_ventilation_ml = minute_ventilation..mL.min...api.datatype.36.., 
     cadence = cadence..spm...api.datatype.53.., 
     activity_g = activity..g...api.datatype.49..,
     minute_ventilation_l = minute_ventilation..L.min...api.datatype.36..
  )
```

### Delete useless columns

```{r}
hexo_data <- subset(hexo_data, select = -sleep_position..NA...api.datatype.270..)
hexo_data <- subset(hexo_data, select = -X)
hexo_data <- subset(hexo_data, select = -time)
```

### Ventilation for Lyon

Previous analysis showed that ventilation for Lyon was missing for some reason. Suspicion is that only L was stored and ml for some reason. Check that suspicion now. 

```{r}
city_vent_na <- hexo_data %>%
          group_by(city, participants) %>%
            summarize(
                vent_ml = mean(minute_ventilation_ml),
                vent_L = mean(minute_ventilation_l)
                )
print(city_vent_na)

city_vent <- hexo_data %>%
          group_by(city, participants) %>%
            summarize(
                vent_ml = mean(minute_ventilation_ml, na.rm = TRUE),
                vent_L = mean(minute_ventilation_l, na.rm = TRUE)
                )
print(city_vent)
```

Yes. Looks like Lyon was only stored as a L and not mL. Also, the litres for Lyon are not litres. They look more like ml. I'm going to assume that everything is ml and check back on this later. 

### Fixing ventilation for Lyon 
```{r}
hexo_data <- hexo_data %>% 
              replace_na(list(minute_ventilation_ml = 0, 
                              minute_ventilation_l = 0)
                         )

hexo_data$minute_ventilation_ml_city <- hexo_data$minute_ventilation_ml + hexo_data$minute_ventilation_l

city_vent_test <- hexo_data %>%
          group_by(city, participants) %>%
            summarize(
                vent_ml_true = mean(minute_ventilation_ml_city, na.rm = TRUE),
                vent_ml_false = mean(minute_ventilation_ml_city, na.rm = FALSE)
                )
city_vent_test 

hexo_data$minute_ventilation_l_city <- hexo_data$minute_ventilation_ml_city/1000
```

### Add participant individual level data

```{r}
table(hexo_data$participants)

participants <- read_csv("Hexoskin_participants.csv")

table(participants$participants)

hexo_data_complete <- inner_join(hexo_data, participants, by = "participants")

write.csv(hexo_data_complete, "hexo_data.csv") ## Saving so that I do not break my computer when working on this data! 
```

SP is present in the participant data but not in the HexoSkin data. Need to check to confirm that is correct. 

### Make new minute level variable (minute) in hexo_data_complete

```{r}
hexo_data_complete$a <- ymd_hm(hexo_data_complete$DATETIME) # Produces NA when format is not "%m/%d/%Y"
hexo_data_complete$b <- ymd_hms(hexo_data_complete$DATETIME) # Produces NA when format is not "%d.%m.%Y"
hexo_data_complete$a[is.na(hexo_data_complete$a)] <- hexo_data_complete$b[!is.na(hexo_data_complete$b)] # Combine both while keeping their ranks

#hexo_data_complete$newDATETIME <- ymd_hms(hexo_data_complete$DATETIME)
hexo_data_complete$minute <- floor_date(hexo_data_complete$a, "minute")
```

### Group by minute

```{r}
hxd <- hexo_data_complete %>%
                  group_by(participants, minute) %>% 
                    summarize(
                      rr = mean(respiratory_rate, na.rm = TRUE),
                      hr = mean(heart_rate, na.rm = TRUE),
                      ve_L = mean((minute_ventilation_l_city), na.rm = TRUE), 
                      g = sum(activity_g, na.rm = TRUE),
                      age = first(age),
                      city = first(city), 
                      height = first(height_cm), 
                      mass = first(mass_kg), 
                      sex = first(gender)
                    )
```

### Complete Cases

```{r}
hxd <- hxd %>% drop_na()
```

1 case removed to due to missing data. 

### Descriptives (for heart rate, breathing) psych package describe function

```{r}
describe(hxd$hr)

ggplot(hxd, aes(hr)) + 
    geom_histogram()
```

Removing HR values less than 50 as they are implausible for this type of analysis
```{r}
hxd <- subset(hxd, hr >= 50)
```

Find heart rates

```{r}
max_hr <- hxd %>%
  group_by(participants) %>%
  summarise(
         maxhr = max(hr)
  )
```

Make max HR col, convert to percent O2, make VO% relative as absolute value, bind
TO DO 220-age and age/gender
Add DOB column

```{r}
hxd_VO <- hxd %>%
  group_by(participants) %>%
    mutate(
      hrmax = max(hr))
#na.omit(hxd_VO$hrmax)
#hxd_VO$percentmaxhr <- NULL

hxd_VO <- hxd_VO %>%
   mutate(
     percentmaxhr = ((hr/hrmax)*100),
     VO = ((percentmaxhr-37.182)/0.6483),
     hr_max_220 = (220 - age)
   )

hexo_data <- bind_rows(hxd_VO)
```

Categorize VO by METs

```{r}
hexo_data <- hexo_data %>%
  mutate(MET = case_when(
  VO <= 3.5 ~ 1, 
  VO > 3.50000001 & VO < 7 ~ 2,
  VO > 7.00000001 & VO < 10.5 ~ 3,
  VO > 10.500000001 & VO < 14.0 ~ 4,
  VO > 14.000000001 & VO < 17.5 ~ 5,
  VO > 17.500000001 & VO < 21.0 ~ 6,
  VO > 21.000000001 & VO < 24.5 ~ 7, 
  VO > 24.500000001 & VO < 28.0 ~ 8, 
  VO > 28.000000001 & VO < 31.5 ~ 9, 
  VO > 31.500000001 & VO < 35.0 ~ 10, 
  VO > 35.000000001 & VO < 38.5 ~ 11, 
  VO > 38.500000001 & VO < 42.0 ~ 12, 
  VO > 42.000000001 & VO < 45.5 ~ 13, 
  VO > 45.500000001 & VO < 49.0 ~ 14, 
  VO > 49.000000001 & VO < 52.5 ~ 15,
  VO > 52.500000001 & VO < 56.0 ~ 16,
  VO > 56.000000001 & VO < 59.5 ~ 17,
  VO > 59.500000001 & VO < 63.0 ~ 18,
  VO > 63.000000001 & VO < 66.5 ~ 19, 
  VO > 66.500000001 & VO < 70.0 ~ 20,
  VO > 70.000000001 & VO < 73.5 ~ 21, 
  VO > 73.500000001 & VO < 77.0 ~ 22,
  VO > 77.000000001 & VO < 80.5 ~ 23,
  VO > 80.500000001 & VO < 84.0 ~ 24, 
  VO > 84.000000001 & VO < 87.5 ~ 25, 
  VO >= 87.50000001 ~ 26
  ))
```

```{r}
describe(hexo_data$MET)
```

MET histos

### scatterplot for predictor and outcome

```{r}
cor_bar <- ggplot(hexo_data, aes(x = MET, y = ve_L)) + 
                  geom_point(size = 0.05) + 
  geom_smooth(method='lm', formula = y ~ x) + 
  ylab("MInute Ventilation") + 
  xlab("Metabolic Equivelents of Task") + 
  ggtitle("Cycle commuting intensity vs Ventilation") +
  facet_wrap(~ city)
plot(cor_bar)
```

### Histogram of all data

```{r}
ggplot(hexo_data, aes(MET)) +
  geom_histogram(aes(y = ..density..), fill = "cornflowerblue", colour = "black", size = 0.5) +
  geom_density(aes(y = ..density..), colour = "darkred", size = 0.8) + 
  ylab("MET density") + 
  xlab("METs")
```

### Writing data for analysis
```{r}
write_csv(hexo_data, "hexo_data_analysis.csv")
```